### 2024-03-11 paper/paper.qmd
added paper template and simulate section.

---
title: "Investigating the correlation between a vote for Biden or Trump based on a persons education, gender, race and age."
author: 
  - Fares Alkorani
  - Owais Zahid
thanks: "Code and data are available at: https://github.com/FFFiend/linear_model_investigation."
date: today
date-format: long
abstract: "In this paper, we discuss the correlation between various statistics of the respondents to the CES2020 election survey, and whether they voted for Trump or Biden. The statistics are: age, education, gender and race."
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Simulation
We first start with simulating the dataset, and recall that we are concerned with each voter's education level, gender, race and age. Our chosen sample size is 8000, and the categories for education and race have been named akin to the categories in the dataset. Thus, the header of the table obtained for the simulation is shown below:
```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false
# setting seed, prime number :P
set.seed(71)
library(dplyr)
# sample size 8000
num_obs <- 8000

# simulating the actual data
# using 8 categories for races as in the dataset, and ages 0-100.
us_political_preferences <- tibble(
  education = sample(0:4, size = num_obs, replace = TRUE),
  gender = sample(0:1, size = num_obs, replace = TRUE),
  race = sample(1:8, size = num_obs, replace = TRUE),
  age = sample(0:100, size = num_obs, replace=TRUE),
  support_prob = ((education + gender + race + age) / 7),
) |>
  mutate(
    supports_biden = if_else(runif(n = num_obs) < support_prob, "yes", "no"),
    education = case_when(
      education == 0 ~ "< High school",
      education == 1 ~ "High school",
      education == 2 ~ "Some college",
      education == 3 ~ "College",
      education == 4 ~ "Post-grad"
    ),
    gender = if_else(gender == 0, "Male", "Female"),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern",
      race == 7 ~ "Two or more races",
      race == 8 ~ "Other"
    )
  ) |>
  select(-support_prob, supports_biden, gender, education, race, age)

# preview the simulated data
cat("Table 1. Header for the political preferences dataset.")
head(us_political_preferences)
```

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

## Diagnostics

\newpage

# References